The Strong Gravitational Lens Finding Challenge

### 2017-09-07 paper.tex
Edit

\begin{table*}
\centering
\begin{tabular}{rlll}
  \hline
 & Name & type & authors \\ 
  \hline
  1 & AstrOmatic & Space-Based & Bertin   \\ 
  2 & GAHEC IRAP & Space-Based & Cabanac  \\ 
  3 & CAS Swinburne Melb & Ground-Based & Jacobs   \\ 
  4 & ALL-star & Ground-Based & Avestruz, N.Li \& Lightman  \\ 
  5 & Manchester1 & Space-Based & Jackson \& Tagore  \\ 
  6 & CMU-DeepLens-Resnet-Voting & Space-Based & Ma, Lanusse \& C. Li  \\ 
  7 & Manchester SVM & Ground-Based & Hartley \& Flamary  \\ 
  8 & CMU-DeepLens-ResNet & Space-Based & Francois Lanusse, Ma, C. Li \& Ravanbakhsh \\ 
  9 & CMU-DeepLens-Resnet-Voting & Ground-Based & Ma, Lanusse \& C. Li   \\ 
  10 & YattaLensLite & Space-Based &  Sonnenfeld  \\ 
  11 & NeuralNet2 & Space-Based &  Davies  \\ 
  12 & CAST & Ground-Based &  Roque De Bom, Valent\'{\i}n \&  Makler  \\ 
  13 & CMU-DeepLens-ResNet-ground3 & Ground-Based &  Lanusse, Ma, Ravanbakhsh \& C. Li \\ 
  14 & GAMOCLASS & Space-Based & Marc Huertas-Company, Tuccillo, Velasco-Forero \& Decenci\~{a}re \\ 
  15 & LASTRO EPFL (CNN) & Space-Based & Geiger, Sch\"{a}fer \& Kneib \\ 
  16 & Manchester SVM & Space-Based &  Hartley  \& Flamary\\ 
  17 & CMU-DeepLens-ResNet-aug & Space-Based & Ma,  Lanusse, Ravanbakhsh \& C. Li \\ 
  18 & LASTRO EPFL & Ground-Based & Geiger, Sch\"{a}fer \& Kneib   \\ 
  19 & CAST & Space-Based & Bom, Valent\'{\i}n \& Makler  \\ 
  20 & AstrOmatic & Ground-Based & Bertin  \\ 
  21 & All-now & Space-Based & Avestruz, N. Li \& Lightman   \\ 
  22 & Manchester-NA2 & Ground-Based & Jackson \& Tagore  \\ 
  23 & YattaLensLite & Ground-Based & Sonnenfeld   \\ 
  24 & Kapteyn Resnet & Space-Based & Petrillo, Koopmans, Kleijn, Tortora \& Vernardos \\ 
   \hline
\end{tabular}
\caption{Entries to the challenges.}
\label{table:entries}
\end{table*}

@@

This section contains short descriptions of the lens finding methods that were used in 
the challenge.  Each subsection refers to a team which gave a separate entry.

{\color{red} MM: May be we could separate the methods into three categories: machine-learning, arc-finders, visual inspection and group them accordingly}

@@

All images (a total of 100000) were examined for each of the space- and
ground-based datasets. This was done by two observers; AT examined 30000
images in each case and NJ examined 70000. Observation was carried out
over a 48-hour period, at the rate of 5000/hr (NJ) and 2500/hr (AT). 
The overall results, in terms of ROC curves, were very similar for 
both observers. The space-based challenge produced areas of 0.800 and 
0.812 for NJ and AT respectively, and the ground-based challenge yielded 
0.891 and 0.884.

The Python scripts used for manual examination of multiple images are available
onGitHub\footnote{https://github.com/nealjackson/bigeye} and are described in more detail
in \citet{hartley2017support}. For one-colour data such as
the space-based training set, the images are individually colour-scaled using
square-root scaling. The bright limit of the colour-scale is determined
from the pixel values in a rectangle comprising the inner ninth of the
image area, with the limit being chosen as the $n$th centile of the pixel
values in this area. Values between $n=95$ and $n=98$ give optimum results,
judging by experiments on the training set. The number of images in each
grid was also optimised using the training set, with 16$\times$8 or 
8$\times$4 giving good results on one-colour data. For three-colour data,
such as the ground-based challenge data, the individual bands for each 
object are colour-scaled and then combined into an RGB image. In this case
8$\times$4 grids were used for examination, due to the generally lower 
resolution of the images. The script also allows the user to adjust the
colour-scale in real time when examining and marking images, and records
the image name corresponding to the image within which the cursor resides
at the time any key is pressed, together with the key.

@@

\subsection{Arc-Finders}
\subsubsection{ IRAP (Cabanac)}

@@

Arcfinder \citep{2006astro.ph..6757A,2007A&A...461..813C,2012ApJ...749...38M} illustrated Figure \ref{fig:Cabanac}, is a fast linear method that computes a pixel wise elongation parameter (ratio of first-order moments in a n-pix window oriented in proper reference frame) for all pixels of mexican-hat-smoothed FITS images. Arcfinder then extracts contiguous pixels above a given background and computes the candidate arcs length, width, area, radius of curvature and peak surface brightness. A final thresholding is set to maximize purity over completeness on a few typical arcs of the dataset.
For the current SL challenge, arcfinder was tuned to detect long and narrow arcs, and was optimized on a subset of 1000 simulated images with a grid covering a range of elongation windows and arc areas.  A python wrapper allows users to change parameters in a flexible way and run the arcfinder C code as a linux line command. Arcfinder took a couple of hours to run on the entire dataset with some overheads due to the dataset format. The code is publicly available at https://github.com/rcabanac/arcfinder.

@@

An alternative approach was to design a set of Gabor filters to be applied to each sample. The Gabor kernel is described by a sinusoidal function multiplied by a Gaussian envelope. We discard the imaginary part of the function to leave, in two-dimensional space:

\begin{equation}
G_c[i,j]=Be^{-\frac{(i^2+j^2)}{2\sigma^2}} \mathrm{cos}\left(\frac{2\pi}{\lambda} (i\, \mathrm{cos} \, \theta + j\, \mathrm{sin} \,\theta)\right),
\end{equation}

where harmonic wavelength $\lambda$, Gaussian spread  $\sigma$ and orientation $\theta$ define the operation performed on each point $i,j$ in an image. Such a kernel is a popular image processing choice for edge detection and texture classification \citep[e.g.][]{Feichtinger98a,Springer-verlag97computationalmodels} and is thought to mimic some image processing functions of the mammalian brain \citep{Jones1233}.

@@

%-------------------------------------------------%
\subsubsection{ALL Methodology (Avestruz, Li, Lightman}
%-------------------------------------------------%

@@

\subsection{Convolutional Neural Networks ((Lanusse)}
Since Convolutional Neural Networks (CNNs) are central to many of the methods that will be described later a brief general description of them will be useful.
A CNN \citep{Lecun1998} is a multi-layer feed-forward neural network model particularly well suited for processing natural images. With the very recent advances of the \textit{Deep Learning} framework \citep{Lecun2015}, models based on CNN architectures have reached or even surpassed human accuracy in image classification tasks \citep{He2015a}.

The fundamental building block of a CNN is the \textit{convolutional layer}. This element applies a set of convolution filters on an input image to produce a series of so-called \textit{feature maps}. The coefficients of these filters are free parameters that are learned by the model. We will use the notation 
$n \times n - n_c$ to denote a convolutional layer using filters of size $n \times n$ pixels and outputting $n_c$ feature maps.  In typical architectures, the size of these convolution filters is kept small (i.e. 3x3 or 5x5 pixels) to limit the complexity of the model. To capture features on larger scales, CNNs rely on a multi-resolution approach by interleaving convolutional layers with \textit{pooling layers} which apply a downsampling operation to the feature maps. The most common downsampling schemes are the max-pooling and average pooling strategies, which downsample an input image by taking respectively the maximum or average values within a given region (e.g. 2x2 patches for a downsampling of factor 2).

@@

For a binary classification problem such as the one involved in strong lens detection, the training is performed by optimizing the weights of the model as to minimize the \textit{binary cross-entropy} objective:
\begin{equation}
	- \sum_{n=1}^{N} y[n] \log \hat{y}[n] + (1 - y[n])\log(1 - \hat{y}[n]) \;,
\end{equation}
where $N$ is the number of training instances, $y \in \{0,1\}$ is the true class of the image and $\hat{y} \in [0,1]$ is the class probability predicted by the model. This optimization is performed by a Stochastic Gradient Descent (SGD) algorithm which updates the model iteratively by taking small gradient steps over randomly selected subsamples of the training set (so called \textit{mini-batches}). All the CNN-based methods presented in this work rely on the ADAM \citep{Kingma_2014} algorithm, an accelerated variant of SGD algorithm.

@@

@@

The lens detector is based on a convolutional neural network (CNN), trained with the provided training datasets. The CNN is implemented in Python, using the TensorFlow framework\footnote{http://www.tensorflow.org/}. Both ground multichannel and space monochannel image classifiers have the exact same CNN architecture.

{\color{red} (MM: I think that the co-authors should provide a more detailed description of the algorithms. For example, they should quickly explain what is a CNN and what are its general components (i.e. what kind of layers, activation functions, etc.)} 
{\color{blue} RBM: I moved the general description of the CNN to the beginning of this section.}

The network itself consists of three convolutional layers ($11\times 11\times 32$, $5\times 5 \times 64$ and $3\times 3\times 64$), followed by two fully-connected layers ($256$ and $64$ neurons) and an output softmax layer {\color{red} MM: the numbers make little sense without a description of what the mean -- see my previous comment}. The first five layers use the ELU (Exponential Linear Unit) activation function \citep{2015arXiv151107289C}, which in our tests led to significantly faster convergence compared to ReLU and even SoftPlus activation. Dropout regularization \citep{2012arXiv1207.0580H,JMLR:v15:srivastava14a} is applied to both convolutional and fully connected layers, with ``keep'' probabilities $p=2/3$ and $p=1/2$, respectively.

Prior to entering the first convolutional layer, input image data are rescaled and dynamic-range compressed with function $f(x) =
\mathrm{arcsinh} (10^{11} x)$, and bad pixels are simply set to 0.
Data augmentation is performed in the form of random up-down and left-right image flipping, plus $k\pi/2$ rotations, where $k$ is a random integer in the $[0,3]$ range. Additionally, a small rotation with random angle $\theta$ is applied, involving bicubic image resampling. $\theta$ follows a Gaussian distribution with mean $\mu=0$ and standard deviation $\sigma_{\theta}=5^{\circ}$. No attempt was made to generate and randomize bad pixel masks in the data augmentation process.

@@

We used a convolutional neural network \citep{Fukushima1980,Lecun1998} with a simple architecture of 12 layers (inspired by \citep{symmetry}), see table \ref{tab:architecture}.
To avoid the problem of the data flow distribution getting out of the comfort zone of the activation functions (Internal Covariate Shift), we used a mix of normalization propagation \citep{norm_prop} (without the constraint on the weights but a proper initialization) and batch normalization \citep{batch_norm} (slowly disabled over the iterations).
As activation function, we used a scaled and shifted ReLU (Rectifier Linear Unit), 
\begin{equation} \label{eq:relu}
    \frac{1}{\sqrt{\pi-1}} (\sqrt{2 \pi} \max(0,x) - 1),
\end{equation}
to satisfy the properties required by the normalization propagation.
In our implementation of batch normalization, instead of using only the moments computed over the current batch, we used a low-pass filter ($\bar\mu_i \longleftarrow (1-\eta) \; \bar\mu_{i-1} + \eta \; \mu_i(\text{batch})$, $\eta$ set to $1$ at the beginning and decay with the iterations).
For the training, the 20k provided images were split into two sets, 17k for training and 3k for validation.
Each iteration of the gradient descent (more precisely \textit{Adam} \citep{adam}) minimizes the cross entropy, 
\begin{equation} \label{eq:xent}
    \left\{
    \begin{array}{ll}
        - \log(p)   & \text{if the image is a true lens} \\
        - \log(1-p) & \text{if the image is a nonlens}
    \end{array}
    \right.,
\end{equation}
where $p$ is the output of the neural network, computed over a batch of 30 images, 15 lenses and 15 nonlenses, picked from the training set.
Making batches of 30 images adds noise to the gradient makes the iterations faster.

@@

Our algorithm for classification of strong lensed galaxies was developed as a convolutional neural network (CNN) used for binary classification task.
Our code was trained and tested on the space-based dataset (dataset 0) provided for the Strong lensing challenge, comprising 20,000 stamps of single band simulated galaxies. We used the full dataset in the proportion of 4/5 for training and 1/5 for validation. The training images were labelled with 1 if showing strong lensing and 0 otherwise. Our CNN gives as output a probability [0,1] of the input image being a strongly lensed system.
The final architecture of our model is illustrated in Fig. \ref{Fig1_gamoclass}. The input image (101x101 pixels) is first processed by a 2D convolution layer with a 3x3 filter size, then subsampled by a 3x3 max pooling layer. Other two identical units follow, with a growing dimensionality of the output space (depth, i.e.output number of filters) in the convolution, for a total of 3 convolutional layers and 3 max pooling layers. Each of these convolutional layers is followed by a rectified linear unit (ReLU) step. The output of these units is then processed through a single fully-connected layer follower by a dropout layer, and, finally, by a one-neuron fully connected layer with sigmoid activation function. For the classification problem we used the binary cross-entropy cost function and for weights updates we used an adaptive moment estimation (ADAM, \citealt{Kingma_2014}) optimization method. The use of the ADAM optimizer improves the learning rate, and previous tests with stochastic gradient descent (SGD) lead to worse results on the classification of the validation sample.
In order to increase the size of the training set and make the model invariant to specific transformations, we perform several data augmentation steps:
\begin{itemize}
\item Random rotation of the image in the range [0, 180$\degree$], using a reflection fill mode to keep constant the size of the images.
\item The images were randomly shifted of 0.02 times the total width of the image.
\item Random horizontal and vertical flips of the images
\end{itemize}
During the training we initialize the weights of our model with random normal values and we warm up the training \citet{Huang_2016} of the CNN for 25 epochs, using an exponential decay rate (1e-6) and a staring learning rate of α = 0.001. Then the network is trained using an early stopping method, and for a maximum number of 300 epochs. The early stopping method is an effective method to prevent overfitting and consists in stopping the training if a monitored quantity does not improve for a fixed number (called \textit{patience}) of training epochs. The quantity that we monitored was the \textit{accuracy} of the classification of the validation sample. The best architecture was trained over 220 epochs with a parameter of patience equal to 20.
We implemented our code the Keras framework \citet{Chollet_2015} on top of Theano \citet{Bastien_2012}.
Our architecture converges with a classification accuracy of 91\% on the validation sample. We further evaluated the performance of our classifier calculating the ROC curve of the classifier, i.e. the True Positive Rate (TPR) against the False Positive Rate. As shown in Fig. \ref{Fig2_gamoclass}, we reach a TPR higher than the 90\% with a FPR < 8\%.

@@

%-------------------------------------------------%
\subsubsection{CAST (Bom, Valent\'in, Makler)}
%-------------------------------------------------%
[Note: we are placing this section here, but we agree it is good to group the methods. Perhaps machine learning based methods could go on a separate section and a unified description can be given to CNN-based methods, as most follow under this category.]

@@

As mentioned above, the CNN used can take 3 images as input. Thus, if we want to use the information on the $4$ available bands, we will have to either combine $2$ of the $4$ bands to end up with $3$ bands for a single RGB CNN (configuration I below) or we use multiple CNNs to use the information available (configuration II to VI below). To combine the outputs of several CNNs we use a Support Vector Machine \citep[hereafter SVM; see e.g., ][]{rebentrost2014quantum} also implemented in Matlab. The SVM is used to combine the outputs $p_i$ of the several CNN (configurations II, III, IV and VI). Instead of using only $p_i$ as inputs to the SVM we also tested providing the SVM with image features obtained by the CNN (feature maps, configuration V) as inputs. In all cases we tested with and without the preprocessing

@@

Much like conventional CNNs, resnets are based on convolutional and pooling layers. However, resnet differ from CNNs by the introduction of so-called shortcut connections bypassing blocks of several convolutional layers. As a result, instead of learning the full mapping from their input to their output these residual blocks only have to learn the difference to the identity. In practice, this difference allows residual networks to be trained even for very deep models. For a more thorough description of this architecture, we refer the interested reader to Section~2.3 of \citet{Lanusse2017}.

Our baseline model is composed of a first 7x7-32 convolutional layer which can accommodate either single-band or multi-band images. The rest of the model is composed of 5 successive blocks, each block being made of 3 resnet units (specifically, pre-activated bottleneck residual units \citep{He2016}). At each block, the signal is downsampled by a factor 2 and the number of feature maps is in turn multiplied by 2. The model is topped by an average-pooling layer followed by a single fully-connected sigmoid layer with a single output. Apart from the final layer, we use the ELU activation throughout. The weights of the model are initialized using random normal values, following the strategy advocated in \citet{He2015a}.

@@

We adopt a minimal pre-processing pre-processing strategy for the input images, removing the mean image and normalising by the noise standard deviation $\sigma$ in each band, these statistic being evaluated over the whole training set. In addition, we clip extreme values above $250 \sigma$ to limit the dynamic range of the input. Bad-pixels are simply set to 0 after this pre-processing step.

@@

Our model is implemented using the \texttt{Theano}\footnote{\url{http://deeplearning.net/software/theano/}} and \texttt{Lasagne}\footnote{\url{https://github.com/Lasagne/Lasagne}} libraries. On an Nvidia Titan Xp GPU, our full training procedure requires approximately 6 hours on the ground-based challenge, but classification of the whole testing set is performed in a couple of minutes. Finally, in the interest of reproducible research, our code is made publicly available at
\begin{center}
\url{https://github.com/McWilliamsCenter} 
\end{center}
This repository also contains a notebook detailing how to reproduce our challenge submission.

\subsubsection{Kapteyn Resnet (Petrillo, Tortora, Verdoes Kleijn, Koopmans) }
Our lens-finder is based on a Convolutional Neural Network (CNN), following the strategy adopted recently in \citep{2017arXiv170207675P}.

@@

\red{breakups Manchester between Neil and Amit}

>>>>

\begin{table*}
\centering
\begin{tabular}{rlll}
  \hline
 & Name & type & authors \\ 
  \hline
  1 & AstrOmatic & Space-Based & Bertin   \\ 
  2 & GAHEC IRAP & Space-Based & Cabanac  \\ 
  3 & \red{CAS Swinburne Melb} & Ground-Based & Jacobs   \\ 
  4 & ALL-star & Ground-Based & Avestruz, N.Li \& Lightman  \\ 
  5 & Manchester1 & Space-Based & Jackson \& Tagore  \\ 
  6 & CMU-DeepLens-Resnet-Voting & Space-Based & Ma, Lanusse \& C. Li  \\ 
  7 & Manchester SVM & Ground-Based & Hartley \& Flamary  \\ 
  8 & CMU-DeepLens-ResNet & Space-Based & Francois Lanusse, Ma, C. Li \& Ravanbakhsh \\ 
  9 & CMU-DeepLens-Resnet-Voting & Ground-Based & Ma, Lanusse \& C. Li   \\ 
  10 & YattaLensLite & Space-Based &  Sonnenfeld  \\ 
  11 & \red{NeuralNet2} & Space-Based &  Davies  \\ 
  12 & CAST & Ground-Based &  Roque De Bom, Valent\'{\i}n \&  Makler  \\ 
  13 & CMU-DeepLens-ResNet-ground3 & Ground-Based &  Lanusse, Ma, Ravanbakhsh \& C. Li \\ 
  14 & GAMOCLASS & Space-Based & Marc Huertas-Company, Tuccillo, Velasco-Forero \& Decenci\~{a}re \\ 
  15 & LASTRO EPFL (CNN) & Space-Based & Geiger, Sch\"{a}fer \& Kneib \\ 
  16 & Manchester SVM & Space-Based &  Hartley  \& Flamary\\ 
  17 & CMU-DeepLens-ResNet-aug & Space-Based & Ma,  Lanusse, Ravanbakhsh \& C. Li \\ 
  18 & LASTRO EPFL & Ground-Based & Geiger, Sch\"{a}fer \& Kneib   \\ 
  19 & CAST & Space-Based & Bom, Valent\'{\i}n \& Makler  \\ 
  20 & AstrOmatic & Ground-Based & Bertin  \\ 
  21 & All-now & Space-Based & Avestruz, N. Li \& Lightman   \\ 
  22 & Manchester-NA2 & Ground-Based & Jackson \& Tagore  \\ 
  23 & YattaLensLite & Ground-Based & Sonnenfeld   \\ 
  24 & Kapteyn Resnet & Space-Based & Petrillo, Koopmans, Kleijn, Tortora \& Vernardos \\ 
   \hline
\end{tabular}
\caption{Entries to the challenges.}
\label{table:entries}
\end{table*}

@@

This section contains short descriptions of the lens finding methods that were used in 
the challenge.  Each subsection refers to a team which gave a separate entry.  We have grouped the 
methods into four categories according to the type of method used.

@@

All images (a total of 100000) were examined for each of the space- and
ground-based datasets. This was done by two observers; AT (Amit Tagore) examined 30000
images in each case and NJ (Neal Jackson) examined 70000. Observation was carried out
over a 48-hour period, at the rate of 5000/hr (NJ) and 2500/hr (AT). 
The overall results, in terms of ROC curves, were very similar for 
both observers. The space-based challenge produced areas of 0.800 and 
0.812 for NJ and AT respectively, and the ground-based challenge yielded 
0.891 and 0.884.

The Python scripts used for manual examination of multiple images are available
on GitHub\footnote{https://github.com/nealjackson/bigeye} and are described in more detail
in \citet{hartley2017support}. For one-colour data such as
the space-based training set, the images are individually colour-scaled using
square-root scaling. The bright limit of the colour-scale is determined
from the pixel values in a rectangle comprising the inner ninth of the
image area, with the limit being chosen as the $n$th centile of the pixel
values in this area. Values between $n=95$ and $n=98$ give optimum results,
judging by experiments on the training set. The number of images in each
grid was also optimised using the training set, with 16$\times$8 or 
8$\times$4 giving good results on one-colour data. For three-colour data,
such as the ground-based challenge data, the individual bands for each 
object are colour-scaled and then combined into an RGB image. In this case
8$\times$4 grids were used for examination, due to the generally lower 
resolution of the images. The script also allows the user to adjust the
colour-scale in real time when examining and marking images, and records
the image name corresponding to the image within which the cursor resides
at the time any key is pressed, together with the key.

@@

\subsection{Arc-Finders}

These methods seek to identify gravitationally lensed arcs and differentiate between them and 
other objects such as spiral arms and edge on spirals using their width, color, curvature and other pre-selected criterion.

\subsubsection{ IRAP (Cabanac)}

@@

Arcfinder \citep{2006astro.ph..6757A,2007A&A...461..813C,2012ApJ...749...38M} illustrated in Figure \ref{fig:Cabanac}, is a fast linear method that computes a pixel wise elongation parameter (ratio of first-order moments in a n-pix window oriented in proper reference frame) for all pixels of mexican-hat-smoothed FITS images. Arcfinder then extracts contiguous pixels above a given background and computes the candidate arc's length, width, area, radius of curvature and peak surface brightness. A final thresholding is set to maximize purity over completeness on a few typical arcs of the dataset.
For the current SL challenge, arcfinder was tuned to detect long and narrow arcs, and was optimized on a subset of 1000 simulated images with a grid covering a range of elongation windows and arc areas.  A python wrapper allows users to change parameters in a flexible way and run the arcfinder C code as a linux line command. Arcfinder took a couple of hours to run on the entire dataset with some overheads due to the dataset format. The code is publicly available at https://github.com/rcabanac/arcfinder.

@@

An alternative approach was to design a set of Gabor filters to be applied to each sample. The Gabor kernel is described by a sinusoidal function multiplied by a Gaussian envelope. We discard the imaginary part of the function to leave, in two-dimensional space:
\begin{equation}
G_c[i,j]=Be^{-\frac{(i^2+j^2)}{2\sigma^2}} \mathrm{cos}\left(\frac{2\pi}{\lambda} (i\, \mathrm{cos} \, \theta + j\, \mathrm{sin} \,\theta)\right),
\end{equation}
where harmonic wavelength $\lambda$, Gaussian spread  $\sigma$ and orientation $\theta$ define the operation performed on each point $i,j$ in an image. Such a kernel is a popular image processing choice for edge detection and texture classification \citep[e.g.][]{Feichtinger98a,Springer-verlag97computationalmodels} and is thought to mimic some image processing functions of the mammalian brain \citep{Jones1233}.

@@

%-------------------------------------------------%
\subsubsection{ALL (Avestruz, Li, Lightman)}
%-------------------------------------------------%

@@

\subsection{Convolutional Neural Networks ((Lanusse)}

Since Convolutional Neural Networks (CNNs) are central to many of the methods that will be described later a brief general description of them will be useful.
A CNN \citep{Lecun1998} is a multi-layer feed-forward neural network model particularly well suited for processing natural images. With the very recent advances of the \textit{Deep Learning} framework \citep{Lecun2015}, models based on CNN architectures have reached or even surpassed human accuracy in image classification tasks \citep{He2015a}.

The fundamental building block of a CNN is the \textit{convolutional layer}. This element applies a set of convolution filters on an input image to produce a series of so-called \textit{feature maps}. The coefficients of these filters are free parameters that are learned by the model. We will use the notation 
$n \times n - n_c$ to denote a convolutional layer using filters of size $n \times n$ pixels and outputting $n_c$ feature maps.\red{(Need to make everyone conform to this notation.)}  In typical architectures, the size of these convolution filters is kept small (i.e. 3x3 or 5x5 pixels) to limit the complexity of the model. To capture features on larger scales, CNNs rely on a multi-resolution approach by interleaving convolutional layers with \textit{pooling layers} which apply a downsampling operation to the feature maps. The most common downsampling schemes are the max-pooling and average pooling strategies, which downsample an input image by taking respectively the maximum or average values within a given region (e.g. 2x2 patches for a downsampling of factor 2).

@@

For a binary classification problem such as the one involved in strong lens detection, the training is performed by optimizing the weights of the model as to minimize the \textit{binary cross-entropy} objective:
\begin{equation}
	- \sum_{n=1}^{N} y[n] \log \hat{y}[n] + (1 - y[n])\log(1 - \hat{y}[n]) \;,
\end{equation}
where $N$ is the number of training instances, $y \in \{0,1\}$ is the true class of the image and $\hat{y} \in [0,1]$ is the class probability predicted by the model. This optimization is usually performed by a Stochastic Gradient Descent (SGD) algorithm which updates the model iteratively by taking small gradient steps over randomly selected subsamples of the training set (so called \textit{mini-batches}). All the CNN-based methods presented in this work rely on the ADAM \citep{Kingma_2014} algorithm, an accelerated variant of SGD algorithm. \red{True?}

@@

\red{Blue terms in following need to be defined.}

@@

The lens detector is based on a CNN, trained with the provided training datasets. The CNN is implemented in Python, using the TensorFlow framework\footnote{http://www.tensorflow.org/}. Both ground multichannel and space monochannel image classifiers have the exact same CNN architecture.

The network itself consists of three convolutional layers ($11\times 11 - 32$, $5\times 5 - 64$ and $3\times 3 - 64$), followed by two fully-connected layers ($256$ and $64$ neurons) and an output \blue{softmax layer}. The first five layers use the ELU (Exponential Linear Unit) \blue{activation function} \citep{2015arXiv151107289C}, which in our tests led to significantly faster convergence compared to ReLU and even SoftPlus activation. Dropout regularization \citep{2012arXiv1207.0580H,JMLR:v15:srivastava14a} is applied to both convolutional and fully connected layers, with ``keep'' probabilities $p=2/3$ and $p=1/2$, respectively.

Prior to entering the first convolutional layer, input image data are rescaled and dynamic-range compressed with function $f(x) =
\mathrm{arcsinh} (10^{11} x)$, and bad pixels are simply set to 0.
\blue{Data augmentation} is performed in the form of random up-down and left-right image flipping, plus $k\pi/2$ rotations, where $k$ is a random integer in the $[0,3]$ range. Additionally, a small rotation with random angle $\theta$ is applied, involving bicubic image resampling. $\theta$ follows a Gaussian distribution with mean $\mu=0$ and standard deviation $\sigma_{\theta}=5^{\circ}$. No attempt was made to generate and randomize bad pixel masks in the data augmentation process.

@@

We used a CNN \citep{Fukushima1980,Lecun1998} with a simple architecture of 12 layers (inspired by \citep{symmetry}), see table \ref{tab:architecture}.
To avoid the problem of the data flow distribution getting out of the comfort zone of the activation functions (\blue{Internal Covariate Shift}), we used a mix of normalization propagation \citep{norm_prop} (without the constraint on the weights but a proper initialization) and batch normalization \citep{batch_norm} (slowly disabled over the iterations).
As activation function, we used a scaled and shifted ReLU (Rectifier Linear Unit), 
\begin{equation} \label{eq:relu}
    \frac{1}{\sqrt{\pi-1}} (\sqrt{2 \pi} \max(0,x) - 1),
\end{equation}
to satisfy the properties required by the normalization propagation.
In our implementation of batch normalization, instead of using only the moments computed over the current batch, we used a low-pass filter ($\bar\mu_i \longleftarrow (1-\eta) \; \bar\mu_{i-1} + \eta \; \mu_i(\text{batch})$, $\eta$ set to $1$ at the beginning and decay with the iterations).
For the training, the 20k provided images were split into two sets, 17k for training and 3k for validation.
Each iteration of the gradient descent (more precisely \textit{Adam} \citep{adam}) minimizes the cross entropy, 
\begin{equation} \label{eq:xent}
    \left\{
    \begin{array}{ll}
        - \log(p)   & \text{if the image is a true lens} \\
        - \log(1-p) & \text{if the image is a nonlens}
    \end{array}
    \right.,
\end{equation}
where $p$ is the output of the neural network, computed over a batch of 30 images, 15 lenses and 15 nonlenses, picked from the training set.
Making batches of 30 images adds noise to the gradient \red{which?} makes the iterations faster.

@@

Our algorithm for classification of strong lensed galaxies was developed as a CNN used for binary classification task.
Our code was trained and tested on the space-based dataset (dataset 0) provided for the Ssrong lensing challenge, comprising 20,000 stamps of single band simulated galaxies. We used the full dataset in the proportion of 4/5 for training and 1/5 for validation. The training images were labelled with 1 if showing strong lensing and 0 otherwise. Our CNN gives as output a probability [0,1] of the input image being a strongly lensed system.
The final architecture of our model is illustrated in Fig. \ref{Fig1_gamoclass}. The input image (101x101 pixels) is first processed by a 2D convolution layer with a 3x3 filter size, then subsampled by a 3x3 max pooling layer.  Another two identical units follow, with a growing dimensionality of the output space (depth, i.e.output number of filters) in the convolution, for a total of 3 convolutional layers and 3 max pooling layers. Each of these convolutional layers is followed by a \blue{rectified linear unit (ReLU) step}. The output of these units is then processed through a single fully-connected layer follower by a dropout layer, and, finally, by a one-neuron fully connected layer with sigmoid activation function. For the classification problem we used the binary cross-entropy cost function and for weights updates we used an adaptive moment estimation (ADAM, \citealt{Kingma_2014}) optimization method. The use of the ADAM optimizer improves the learning rate, and previous tests with stochastic gradient descent (SGD) lead to worse results on the classification of the validation sample.
In order to increase the size of the training set and make the model invariant to specific transformations, we perform several data augmentation steps:
\begin{itemize}
\item Random rotation of the image in the range [0, 180$\degree$], using a \blue{reflection fill mode} to keep constant the size of the images.
\item The images were randomly shifted of 0.02 times the total width of the image.
\item Random horizontal and vertical flips of the images
\end{itemize}
During the training we initialize the weights of our model with random normal values and we warm up the training \citet{Huang_2016} of the CNN for 25 epochs, using an exponential decay rate ($10^{-6}$) and a staring learning rate of α = 0.001. Then the network is trained using an early stopping method, and for a maximum number of 300 epochs. The early stopping method is an effective method to prevent overfitting and consists in stopping the training if a monitored quantity does not improve for a fixed number (called \textit{patience}) of training epochs. The quantity that we monitored was the \textit{accuracy} of the classification of the validation sample. The best architecture was trained over 220 epochs with a parameter of patience equal to 20.
We implemented our code the Keras framework \citep{Chollet_2015} on top of Theano \citep{Bastien_2012}.
Our architecture converges with a classification accuracy of 91\% on the validation sample. We further evaluated the performance of our classifier calculating the ROC curve of the classifier, i.e. the True Positive Rate (TPR) against the False Positive Rate. As shown in Fig. \ref{Fig2_gamoclass}, we reach a TPR higher than the 90\% with a FPR < 8\%.

@@

%-------------------------------------------------%
\subsubsection{CAST (Bom, Valent\'in, Makler)}
%-------------------------------------------------%

@@

As mentioned above, the CNN used can take 3 color images as input. Thus, if we want to use the information on the $4$ available bands, we will have to either combine $2$ of the $4$ bands to end up with $3$ bands for a single RGB CNN (configuration I below) or we use multiple CNNs to use the information available (configuration II to VI below). To combine the outputs of several CNNs we use a Support Vector Machine \citep[hereafter SVM; see e.g., ][]{rebentrost2014quantum} also implemented in Matlab. The SVM is used to combine the outputs $p_i$ of the several CNNs (configurations II, III, IV and VI). Instead of using only $p_i$ as inputs to the SVM we also tested providing the SVM with image features obtained by the CNN (feature maps, configuration V) as inputs. In all cases we tested with and without the preprocessing

@@

Much like conventional CNNs, resnets are based on convolutional and pooling layers. However, resnet differ from CNNs by the introduction of so-called shortcut connections bypassing blocks of several convolutional layers. As a result, instead of learning the full mapping from their input to their output these residual blocks only have to {\color{blue} learn the difference to the identity}\red{(I don't know what you mean by this)}. In practice, this difference allows residual networks to be trained even for very deep models. For a more thorough description of this architecture, we refer the interested reader to Section~2.3 of \citet{Lanusse2017}.

Our baseline model is composed of a first 7x7-32 convolutional layer which can accommodate either single-band or multi-band images. The rest of the model is composed of 5 successive blocks, each block being made of 3 resnet units (specifically, pre-activated bottleneck residual units \citep{He2016}). At each block, the signal is downsampled by a factor 2 and the number of feature maps is in turn multiplied by 2. The model is topped by an average-pooling layer followed by a single fully-connected sigmoid layer with a single output. Apart from the final layer, we use the \blue{ELU} activation throughout. The weights of the model are initialized using random normal values, following the strategy advocated in \citet{He2015a}.

@@

We adopt a minimal pre-processing strategy for the input images, removing the mean image and normalising by the noise standard deviation $\sigma$ in each band, these statistic being evaluated over the whole training set. In addition, we clip extreme values above $250 \,\sigma$ to limit the dynamic range of the input. Bad-pixels are simply set to 0 after this pre-processing step.

@@

Our model is implemented using the \texttt{Theano}\footnote{\url{http://deeplearning.net/software/theano/}} and \texttt{Lasagne}\footnote{\url{https://github.com/Lasagne/Lasagne}} libraries. On an Nvidia Titan Xp GPU, our full training procedure requires approximately 6 hours on the ground-based challenge, but classification of the whole testing set is performed in a couple of minutes. Finally, in the interest of reproducible research, our code is made publicly available on GitHub\footnote{\url{https://github.com/McWilliamsCenter} }.
This repository also contains a notebook detailing how to reproduce our challenge submission.

\subsubsection{Kapteyn Resnet (Petrillo, Tortora, Verdoes Kleijn, Koopmans) }
\red{This is a bit ragged and another submission needs to be analysed.}

Our lens-finder is based on a CNN, following the strategy adopted recently in \citep{2017arXiv170207675P}.

@@

\red{breakups Manchester between Neal and Amit}
